D3064R0: How to Avoid OOTA Without Really Trying

### 2024-02-24 ootavacuous.tex
Split litmus-test appendix into two
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

The earlier examples shown in
Listings~\ref{lst:Simple Reordering}
and~\ref{lst:Simple OOTA}
can be recast as similar \co{herd7} litmus tests, as shown in
Listing~\ref{lst:Simple Reordering as Litmus Test}
on page~\pageref{lst:Simple Reordering as Litmus Test}
and Listing~\ref{lst:Causality Test Case 4}
on page~\pageref{lst:Causality Test Case 4}, respectively.
Other OOTA-related litmus tests may be found in
Appendix~\ref{app:OOTA-Related Litmus Tests}.

@@

\section{OOTA-Related Litmus Tests}
\label{app:OOTA-Related Litmus Tests}

This appendix evaluates a number of OOTA-related litmus tests.
The evaluations are carried out using the \co{herd7} tool and also
using manual analysis.

Appendix~\ref{app:Litmus Tests from “Causality Test Cases"}
evaluates the obligatory litmus tests from the JMM Causality Test
Cases, and
Appendix~\ref{app:Additional Litmus Tests}
evaluates some additional litmus tests that affected our thinking or
otherwise gave us pause.

\subsection{Litmus Tests from “Causality Test Cases"}
\label{app:Litmus Tests from “Causality Test Cases"}

These tests might be two decades old, but they are still quite
relevant.\footnote{
	\url{http://www.cs.umd.edu/~pugh/java/memoryModel/unifiedProposal/testcases.html}.}

\subsubsection{Causality Test Case 1}
\label{app:Causality Test Case 1}

@@

\subsubsection{Causality Test Case 2}
\label{app:Causality Test Case 2}

@@

\subsubsection{Causality Test Case 3}
\label{app:Causality Test Case 3}

@@

\subsubsection{Causality Test Case 4}
\label{app:Causality Test Case 4}

@@

\subsubsection{Causality Test Case 5}
\label{app:Causality Test Case 5}

@@

\subsubsection{Causality Test Case 6}
\label{app:Causality Test Case 6}

@@

\subsubsection{Causality Test Case 7}
\label{app:Causality Test Case 7}

@@

\subsubsection{Causality Test Case 8}
\label{app:Causality Test Case 8}

@@

\subsubsection{Causality Test Case 9}
\label{app:Causality Test Case 9}

@@

\subsubsection{Causality Test Case 9a}
\label{app:Causality Test Case 9a}

@@

\subsubsection{Causality Test Case 10}
\label{app:Causality Test Case 10}

@@

\subsubsection{Causality Test Case 11}
\label{app:Causality Test Case 11}

@@

\subsubsection{Causality Test Case 12}
\label{app:Causality Test Case 12}

@@

\subsubsection{Causality Test Case 13}
\label{app:Causality Test Case 13}

@@

\subsubsection{Causality Test Case 14}
\label{app:Causality Test Case 14}

@@

\subsubsection{Causality Test Case 15}
\label{app:Causality Test Case 15}

@@

\subsubsection{Causality Test Case 16}
\label{app:Causality Test Case 16}

@@

\subsubsection{Causality Test Case 17}
\label{app:Causality Test Case 17}

@@

\subsubsection{Causality Test Case 18}
\label{app:Causality Test Case 18}

@@

\subsubsection{Causality Test Case 19}
\label{app:Causality Test Case 19}

@@

\subsubsection{Causality Test Case 20}
\label{app:Causality Test Case 20}

@@

\subsection{Additional Litmus Tests}
\label{app:Additional Litmus Tests}

These litmus tests helped illuminate important aspects of the problem
of defining sdep and identifying OOTA cycles.

\subsubsection{Simple Reordering as Litmus Test}
\label{app:Simple Reordering as Litmus Test}

@@

\subsubsection{Non-LB Litmus Tests}
\label{app:Non-LB Litmus Tests}

@@

\subsubsection{Multi-Cycle Litmus Tests}
\label{app:Multi-Cycle Litmus Tests}

@@

\subsubsection{Three-Process Version of JMM Causality Test Case 4}
\label{app:Three-Process Version of JMM Causality Test Case 4}

@@

\subsubsection{Why rfe Instead of Tried-And-True rf?}
\label{app:Why rfe Instead of Tried-And-True rf?}

@@

\subsubsection{Inventing Atomic Loads}
\label{app:Inventing Atomic Loads}

@@

\subsubsection{Non-Trivial Semantic Dependencies}
\label{app:Non-Trivial Semantic Dependencies}

@@

\subsubsection{OOTA From Undefined-Behavior Optimizations}
\label{app:OOTA From Undefined-Behavior Optimizations}

@@

\subsubsection{Semantic Dependencies and \co{volatile}}
\label{app:Semantic Dependencies and volatile}

>>>>

The earlier examples shown in
Listings~\ref{lst:Simple Reordering}
and~\ref{lst:Simple OOTA}
can be recast as similar \co{herd7} litmus tests, as shown in
Listing~\ref{lst:Simple Reordering as Litmus Test}
on page~\pageref{lst:Simple Reordering as Litmus Test}
and Listing~\ref{lst:Causality Test Case 4}
on page~\pageref{lst:Causality Test Case 4}, respectively.
Other OOTA-related litmus tests may be found in
Appendix~\ref{app:Litmus Tests from “Causality Test Cases"}
and
Appendix~\ref{app:Additional Litmus Tests}.

@@

\section{Litmus Tests from “Causality Test Cases"}
\label{app:Litmus Tests from “Causality Test Cases"}

These tests might be two decades old, but they are still quite
relevant.\footnote{
	\url{http://www.cs.umd.edu/~pugh/java/memoryModel/unifiedProposal/testcases.html}.}
This appendix evaluates them using the \co{herd7} tool and also using
manual analysis.

\subsection{Causality Test Case 1}
\label{app:Causality Test Case 1}

@@

\subsection{Causality Test Case 2}
\label{app:Causality Test Case 2}

@@

\subsection{Causality Test Case 3}
\label{app:Causality Test Case 3}

@@

\subsection{Causality Test Case 4}
\label{app:Causality Test Case 4}

@@

\subsection{Causality Test Case 5}
\label{app:Causality Test Case 5}

@@

\subsection{Causality Test Case 6}
\label{app:Causality Test Case 6}

@@

\subsection{Causality Test Case 7}
\label{app:Causality Test Case 7}

@@

\subsection{Causality Test Case 8}
\label{app:Causality Test Case 8}

@@

\subsection{Causality Test Case 9}
\label{app:Causality Test Case 9}

@@

\subsection{Causality Test Case 9a}
\label{app:Causality Test Case 9a}

@@

\subsection{Causality Test Case 10}
\label{app:Causality Test Case 10}

@@

\subsection{Causality Test Case 11}
\label{app:Causality Test Case 11}

@@

\subsection{Causality Test Case 12}
\label{app:Causality Test Case 12}

@@

\subsection{Causality Test Case 13}
\label{app:Causality Test Case 13}

@@

\subsection{Causality Test Case 14}
\label{app:Causality Test Case 14}

@@

\subsection{Causality Test Case 15}
\label{app:Causality Test Case 15}

@@

\subsection{Causality Test Case 16}
\label{app:Causality Test Case 16}

@@

\subsection{Causality Test Case 17}
\label{app:Causality Test Case 17}

@@

\subsection{Causality Test Case 18}
\label{app:Causality Test Case 18}

@@

\subsection{Causality Test Case 19}
\label{app:Causality Test Case 19}

@@

\subsection{Causality Test Case 20}
\label{app:Causality Test Case 20}

@@

\section{Additional Litmus Tests}
\label{app:Additional Litmus Tests}

These litmus tests helped illuminate important aspects of the problem
of defining sdep and identifying OOTA cycles.
This appendix evaluates them using the \co{herd7} tool and also using
manual analysis.

\subsection{Simple Reordering as Litmus Test}
\label{app:Simple Reordering as Litmus Test}

@@

\subsection{Non-LB Litmus Tests}
\label{app:Non-LB Litmus Tests}

@@

\subsection{Multi-Cycle Litmus Tests}
\label{app:Multi-Cycle Litmus Tests}

@@

\subsection{Three-Process Version of JMM Causality Test Case 4}
\label{app:Three-Process Version of JMM Causality Test Case 4}

@@

\subsection{Why rfe Instead of Tried-And-True rf?}
\label{app:Why rfe Instead of Tried-And-True rf?}

@@

\subsection{Inventing Atomic Loads}
\label{app:Inventing Atomic Loads}

@@

\subsection{Non-Trivial Semantic Dependencies}
\label{app:Non-Trivial Semantic Dependencies}

@@

\subsection{OOTA From Undefined-Behavior Optimizations}
\label{app:OOTA From Undefined-Behavior Optimizations}

@@

\subsection{Semantic Dependencies and \co{volatile}}
\label{app:Semantic Dependencies and volatile}
